---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo=TRUE}

library(dplyr)
library(lattice)
library(xtable)
library(knitr)
```
## Set global option echo to "TRUE"
```{r setoptions, echo=TRUE}
opts_chunk$set(echo=TRUE)
```


## Loading and preprocessing the data
```{r}
activity<-read.csv("activity/activity.csv")

```
##Convert date type into POSIXct type
```{r}
activity$date <- as.POSIXct(activity$date)
```




# What is mean total number of steps taken per day?


##1. Calculate the total number of steps taken per day, save to *a*
```{r}
activity %>% group_by(date) %>% summarise(total_step = sum(steps, na.rm = TRUE)) ->a
```
##2. Make a histogram of the total number of steps taken each day
##3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
z.plot1<-function(){
        
        hist(a$total_step,col="green",main="Total number of steps taken each day",xlab="Total number of steps")
        abline(v=median(a$total_step), col="red")
        abline(v=mean(a$total_step),col="blue")
        text(mean(a$total_step),25,labels="mean", pos=4, col="blue")  
        text(median(a$total_step),20,labels="median", pos=4, col="red") 
        
        
}
z.plot1()
mean <- mean(a$total_step)
median <- median(a$total_step)

```
The mean of the total number of steps taken per day is `r mean`.
The median of the total number of steps taken per day is `r median`.


# What is the average daily activity pattern?

##1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}

activity %>% group_by(interval) %>% summarise(avg_step = mean(steps, na.rm = TRUE)) ->b

plot(b$interval,b$avg_step, type="l", ylab="Average # of steps", xlab="Interval index", main="5-minute interval")
y<-max(b$avg_step)
x<-subset(b,avg_step==y, select="interval")
points(x,y,col="red",pch=12)
text(x,y,labels="maximum # of steps", pos=4, col="red")  
```


##2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Interval `r x` on average across all the days contains the maximum number of steps `r y`. 


# Imputing missing values

##1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
activity %>% group_by(na=is.na(steps)) %>% summarise(count = n()) ->c
barplot(c$count,names.arg=c("Not NA","NA"),ylim=range(0,nrow(activity)))

na <- subset(c, na,select="count")                     
notna <- subset(c, !na,select="count")    
text(.7,0,labels=notna, pos=3)                                           
text(1.9,0,labels=na, pos=3)

```



##2. Devise a strategy for filling in all of the missing values in the dataset. 

I'll use the average number of steps taken, averaged across all days (y-axis) to impute the missing values at each correspoinding interval.

##3. Create a new dataset dfinal that is equal to the original dataset but with the missing data filled in.



```{r}
merge(activity, b, by="interval") -> d
d[is.na(d["steps"]),]->d1
mutate(d1,steps=avg_step)->d1m
d[!is.na(d["steps"]),]->d2
rbind(d1m,d2)->dfinal
```


##4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}

z.plot2<-function(){
        
        dfinal %>% group_by(date) %>% summarise(total_step = sum(steps, na.rm = TRUE)) ->dfinal2
        
        p2<-hist(dfinal2$total_step,col="green",main="Total number of steps taken each day",xlab="Total number of steps")
        abline(v=median(dfinal2$total_step), col="red")
        abline(v=mean(dfinal2$total_step),col="blue")
        
        text(mean(dfinal2$total_step),25,labels="mean", pos=4, col="blue")  
        text(median(dfinal2$total_step),20,labels="median", pos=4, col="red")  
}

z.plot2()
```

Please note that the red line of median and blue line of mean are overlapped now.That's why the blue line has covered the red line.


##Do these values differ from the estimates from the first part of the assignment? 
```{r}

par(mfrow=c(1,2))
z.plot1()
z.plot2()

```


Yes. The mean and median shift to right and the frequency of the person walks range from 0 to 15000 steps increases. This is because that we impute the NA values with average number of steps at each interval.


##What is the impact of imputing missing data on the estimates of the total daily number of steps?

Increase the number of days that the person walks from 0 to 15000 steps, closer to the average steps walked per day.



# Are there differences in activity patterns between weekdays and weekends?

##1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
mutate(dfinal, weekday=weekdays(date))-> wd
wd[wd["weekday"] == "Saturday" | wd["weekday"] =="Sunday",]->weekend
mutate(weekend, weekday="weekend")-> weekend
wd[wd["weekday"]!="Saturday" & wd["weekday"]!="Sunday",]->weekday
mutate(weekday, weekday="weekday")-> weekday
rbind(weekday,weekend)->wdfinal

wdfinal$weekday <- as.factor(wdfinal$weekday)
```

##2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}

wdfinal %>% group_by(interval, weekday) %>% summarise(avg_step = mean(steps, na.rm = TRUE)) ->wdfinal2


xyplot(avg_step ~ interval | weekday ,wdfinal2, ylab="Average Number of steps", xlab="5-Minute Interval",layout=c(1,2), type=c('l','smooth'), lwd=3)


```


As the plot shows, the individual works out more at weekend than weekdays in general, and less around 8 a.m.at weekend than during weekdays, due to commute to work in the morning.


